---
title: "Lab 00 - Hello IDS!"
author: "adamlaycock"
date: "`r Sys.Date()`"
output: html_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

### Exercises 16-17

We set up the data frame. Edit the data below to include information about your team members.

```{r setup-tibble}
team_data <- tibble(
  name = c("Andrew", "Helen", "Adam", 
           "Max", "Sarah", "Liv"),
  github_username = c("pukklemaster", "helenrmiller", "adamlaycock", 
                      "MaxWellPohlman", "sarahstewart", "oliviaharris1"), 
  colour = c("mediumaquamarine", "thistle1", "forestgreen", 
             "slategray3", "powderblue", "violetred"),
  hobby = c("Knitting", "Crochet", "Guitar", 
            "Climbing", "Swimming", "Netball"),
  favourite_number = c(7, 4, 7, 10, 6, 4),
  hobby_category = c("Arts", "Arts", "Music", "Sports", "Sports", "Sports")
  )

team_data
```

### Exercise 18

We now work on visualising our data. 

```{r bar-plot}
p1 <- ggplot(data = team_data, 
             mapping = aes(x = github_username, 
                           y = favourite_number,
                           fill = colour), 
             ) +
   geom_col() +
   scale_fill_identity()
p1
```


The plot shows a bar chart of each group member's favourite numbers, with the height being proportional to the magnitude of the number.

### Exercises 19

Change the text of the plot title and axis labels.

```{r labelled-bar-plot}
p1 + labs(x = "Favourite Number", 
          y = "Github Username",
          title = "Barchart of favourite number for each GitHub User")
```

labs() is a label function that will assign strings to be displayed on the p1 plot, takes arguments x, y, and title for the x-axis, y-axis, and title respectively.

### Exercise 20

```{r new-bar-plot}
p2 <- ggplot(team_data, aes(x = hobby)) +
  geom_bar()
p2
```

The issue with plotting hobbies is that there would be far too many unique values which would correlate to far too many bars for meaningful visualisation. This could be solved by grouping the hobbies into categories, e.g. knitting, crochet, and painting become an 'Arts' category. This reduces the total number of bars and adds more variety in the bar height.

```{r improved-hobby-plot}
p3 <- ggplot(team_data, aes(x = hobby_category)) + geom_bar()
p3 + labs(x = "Hobby Category", 
          y = "Count",
          title = "Barchart of count of hobby categories for each group member")
```